---
title: "Covariate Regression Analyses"
output: html_document
date: "2023-11-14"
---

Setting up data for analysis and subsetting by group (ie, persisters (at least 2 
consecutive positive samples), coresidents (those who are not persisters but live in the same household as persisters), and non-coresidents (not persisters and do not live in same household as persisters))
```{r}
setwd("/Users/rebeccaperez/Downloads")
load("datafinal.Rdata")

##Subset by subgroup:

persisters <- datafinal[datafinal$PERS == 1,]

pershouses <- unique(persisters$family_code)

pershouse2 <- datafinal[1,]
  for (q in 1:length(pershouses)){
    
    pershouse <- datafinal[datafinal$family_code.x == pershouses[q],]
    
    pershouse2 <- rbind(pershouse2, pershouse)
    
  }

pershouse2 <- pershouse2[-1,]
coresidents <- pershouse2[pershouse2$PERS == 0,]


noncoresidents <- datafinal

for (i in 1:nrow(datafinal)) {
  
for (q in 1:length(pershouses)) { 
  
  if (noncoresidents$family_code.x[i] == pershouses[q]){
    noncoresidents$family_code.y[i] <- NA}
 
}
}

noncoresidents <- noncoresidents[!is.na(noncoresidents$family_code.y),]


datafinal$cores <- 0  ###creating a column to indicate coresident status 

for (i in 1:nrow(datafinal)){
  
  for (q in 1:length(pershouses)){
    
    if(datafinal$family_code.x[i] == pershouses[q]){
      
      datafinal$cores[i] <- 1
      
    if(datafinal$PERS[i] == 1){
      
      datafinal$cores[i] <-0 
    }
    }
  }
}

table(datafinal$cores) #check 


```

Multivariate analysis and plotting of covariates associated with ST131 positive sample 
```{r}

covregmodel1.1<- glm(POS~PERS+medicalcond+incontinent+female + veg + hospital + medicaldevices + age + hometime + famsize + cores, data=datafinal)

summary(covregmodel1.1)

coef(summary(covregmodel1.1))
obj1.1 <- coef(summary(covregmodel1.1))

plotdatameans1.1 <- obj1.1[c(-1),1]
plotdatasds1.1 <- obj1.1[c(-1),2]

plotdataframe <- matrix(nrow = 11, ncol= 5, data = NA)


plotdataframe[,1] <- as.numeric(exp(plotdatameans1.1))
plotdataframe[,2]<- as.numeric(exp(plotdatameans1.1 - (1.96 *plotdatasds1.1)))
plotdataframe[,3]<- as.numeric(exp(plotdatameans1.1 + (1.96 *plotdatasds1.1)))
plotdataframe[,5] <- 1:11
plotdataframe[,4] <-  c("Persistent Carrier*", "Comorbidity", "Incontinence", "Sex (Female)", "Vegetarian", "Hospital Stay in Past Year", "Medical Devices", "Age", ">30 hours per week spent at home", "Family Size", "Coresident (lives with persistent carrier)" )
plotdataframe <- as.data.frame(plotdataframe)
plotdataframe[,1] <- as.numeric(plotdataframe[,1])
plotdataframe[,2] <- as.numeric(plotdataframe[,2])
plotdataframe[,3] <- as.numeric(plotdataframe[,3])

colnames(plotdataframe) <- c("OR", "Lower", "Upper", "Covariate", "Index")

library(extrafont)

g1.0 <- ggplot(plotdataframe, aes(x= OR, y = Covariate)) + geom_errorbar(aes(xmin = Lower, xmax = Upper, width = .5), color="black") +  geom_point(aes(x= OR, y = Covariate), size = 1.8) + geom_vline(xintercept = 1, color = "black", linetype = "dashed") + scale_x_continuous(trans = 'log10', breaks = seq(0, 2, .25), labels = as.character(seq(0,2,.25))) + theme(
  plot.title = element_text(hjust = 0.5),
  plot.subtitle = element_text(hjust = 1),
  plot.caption = element_text(hjust = 0), text =element_text(size=11, family = "Arial"), panel.background = element_rect(fill = "white", color = "white"), panel.grid = element_line(color = "grey")) + xlab(label = "Multivariate Odds Ratio") + ylab(label = "")

g1.0



```

Univariate analysis 

```{r}

univalues <- matrix(data = NA, nrow = 11, ncol = 7)

veg <- glm(POS~veg, data = datafinal, family = "binomial")
obj <- as.data.frame(coef(summary.glm(veg)))


univalues[1,1] <- obj[2,1]
univalues[1,2] <- obj[2,2]


sex <- glm(POS~female, data = datafinal, family = "binomial")
summary(sex)

obj <- as.data.frame(coef(summary.glm(sex)))


univalues[2,1] <- obj[2,1]
univalues[2,2] <- obj[2,2]

pers <- glm(POS~PERS, data = datafinal, family = "binomial")
summary(pers)
obj <- as.data.frame(coef(summary.glm(pers)))


univalues[3,1] <- obj[2,1]
univalues[3,2] <- obj[2,2]

meddevs <- glm(POS~medicaldevices, data = datafinal, family = "binomial")
summary(meddevs)
obj <- as.data.frame(coef(summary.glm(meddevs)))


univalues[4,1] <- obj[2,1]
univalues[4,2] <- obj[2,2]

incont <- glm(POS~incontinent, data = datafinal, family = "binomial")
summary(incont)
obj <- as.data.frame(coef(summary.glm(incont)))


univalues[5,1] <- obj[2,1]
univalues[5,2] <- obj[2,2]

famsize <- glm(POS~famsize, data = datafinal, family = "binomial")
summary(famsize)
obj <- as.data.frame(coef(summary.glm(famsize)))


univalues[6,1] <- obj[2,1]
univalues[6,2] <- obj[2,2]

hosp <- glm(POS~hospital, data = datafinal, family = "binomial")
summary(hosp)
obj <- as.data.frame(coef(summary.glm(hosp)))


univalues[7,1] <- obj[2,1]
univalues[7,2] <- obj[2,2]

cores <- glm(POS~cores, data = datafinal, family = "binomial")
summary(cores)
obj <- as.data.frame(coef(summary.glm(cores)))


univalues[8,1] <- obj[2,1]
univalues[8,2] <- obj[2,2]


comor <- glm(POS~medicalcond, data = datafinal, family = "binomial")
summary(comor)
obj <- as.data.frame(coef(summary.glm(comor)))


univalues[9,1] <- obj[2,1]
univalues[9,2] <- obj[2,2]

age <- glm(POS~age, data = datafinal, family = "binomial")
summary(age)
obj <- as.data.frame(coef(summary.glm(age)))


univalues[10,1] <- obj[2,1]
univalues[10,2] <- obj[2,2]

home <- glm(POS~hometime, data = datafinal, family = "binomial")
summary(home)
obj <- as.data.frame(coef(summary.glm(home)))


univalues[11,1] <- obj[2,1]
univalues[11,2] <- obj[2,2]


univalues[,3] <- exp(univalues[,1])
univalues[,4] <- exp(univalues[,1]-1.96*univalues[,2])
univalues[,5] <- exp(univalues[,1]+1.96*univalues[,2])


univalues <- univalues[,3:6]

colnames(univalues) <- c("OR", "Lower", "Upper", "Index")


Covnames <- as.data.frame(c("Vegetarian", "Sex (Female)", "Persistent Carrier*", "Medical Devices", "Incontinence", "Family Size", "Hospital Stay in Past Year", "Coresident (lives with persistent carrier)", "Comorbidity", "Age", ">30 hours per week spent at home"))

colnames(Covnames) <- "Covariate"

univalues <- cbind(univalues, Covnames)

univalues$Index <- seq(1,11, by = 1)

###remove vegetarian bc infinite 

plotunivalues <- univalues

plotunivalues[1,1:3] <- NA

#####plot 
g1.1 <- ggplot(plotunivalues, aes(x= OR, y = Covariate)) + geom_errorbar(aes(xmin = Lower, xmax = Upper, width = .5), color="black") +  geom_point(aes(x= OR, y = Covariate), size = 1.8) + geom_vline(xintercept = 1, color = "black", linetype = "dashed") + scale_x_continuous(trans = 'log10') + theme(
  plot.title = element_text(hjust = 0.5),
  plot.subtitle = element_text(hjust = 1),
  plot.caption = element_text(hjust = 0), text =element_text(size=11, family = "Arial"), panel.background = element_rect(fill = "white", color = "white"), panel.grid = element_line(color = "grey")) + xlab(label = "Univariate Odds Ratio") + ylab(label = "")

g1.1

```

